A few stickleback suffice to transport adaptive alleles to new lakes

### 2019-01-26 Stickleback_Paper.tex
Fig update
Friday edits, finish on Saturday

Threespine stickleback fish provide a striking example of local adaptation despite recurrent gene flow.
The species is distributed around the Northern hemisphere in both marine and freshwater habitats.
It is thought that these numerous, smaller freshwater populations
have been established ``de novo" from marine fish,
and that a shared freshwater phenotype
is often established using standing genetic variation.
Here we use genealogical simulations to determine the levels of gene flow
that best matches observed patterns of allele sharing among habitats in stickleback, 
and more generally to better understand how gene flow and local adaptation in large metapopulations
determine speed of adaptation and reuse of standing genetic variation.
We find that rapid, repeated adaptation using a shared set of alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.
Low gene flow leads to slow, independent adaptation of distinct habitats, whereas high gene flow leads to large migration load.
We quantify how $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
In addition, we find the origin of many freshwater adapted alleles in the introduced lakes to be propagated from the original generation of marine individuals that had inhabited the lake.
The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.  
\plr{say transporter hypothesis?}
\plr{talk about linkage and haplotypes}

@@

% Jared's general notes for formatting
%
% - ``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)
% - in latex, use ` `words' ' to get the open- and closed-double quotes (not the double-quote character)
% - "citet" for "cite, as part of the text"

The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations. 
While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures. 
What role does genomic variation play in meta-populations with connection to an array of selective pressures?
How could migration-selection balance and ecological speciation actually benefit species as a whole?
A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split \citet{nelson2017ancient,Schrider2017,Barrett2008}.
However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations.
\plr{I like the spirit of the first paragraph, but want to totally rewrite it}

>>>>

Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being contected by recurrent gene flow. These small fish occurin marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases smaller freshwater populations have been established ``de novo'' from marine colonists. Previous results show that a similar freshwater phenotype often evolves in independent populations soon after colonization using largely the same standing genetic variation, even in very distant geographic regions. Here we use genealogical simulations to determine the levels of gene flow that best matches observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and reuse of standing genetic variation. We find that rapid, repeated adaptation using a shared set of alleles that are maintained at low frequency by migration-selection balance in ocean populations occurs over a realistic range of intermediate rates of gene flow that match previous empirical population genomic studies in stickleback. Using these realistic simulations we discovered that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load. A consequence is that population genomic scans for locally adapted alleles are more likely to succeed with higher rates of gene flow. Lastly, we find that the origin of many freshwater adapted alleles in the introduced lakes is the original generation of marine individuals that colonized the lake. Thuses simulations provide deeper context for existing studies of stickleback evolutionary genomics, and provide guidance for future empirical studies in this model. More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.

%\plr{say transporter hypothesis?}
%\plr{talk about linkage and haplotypes}

@@

% Jared's general notes for formatting
% - ``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)
% - in latex, use ` `words' ' to get the open- and closed-double quotes (not the double-quote character)
% - "citet" for "cite, as part of the text"

The sequential fixation of new mutations has long been the canonical model for the genetics of adaptation (\textbf{cite}). While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures. What role does genomic variation play in meta-populations with connection to an array of selective pressures? How could migration-selection balance and ecological speciation actually benefit species as a whole? A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split \citet{nelson2017ancient,Schrider2017,Barrett2008}. However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms. In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations

%\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
